Lyman-alpha spectra from rotating outflows

### 2018-03-27 paper/paper.tex
edit caption
conclusions

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
distributions of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these two
features impact the morphology of the Lyman-$\alpha$ emission line.
One of the most notorious results is that rotation increases the
detected flux at the line's center in comparison with the
non-rotational case. 
Rotation also induces a line broadening and a dependency with the
viewing angle.
We also present a semi-analytic model that modifies the spectra of
non-rotational that manages to reproduce the main features of these
Monte Carlo simulations. 
As an application example we contraint the rotation parameters to the
observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
first a constraint on the rotational velocity of a high redshift
galaxy based on its integrated \lya emission.
\end{abstract}

@@

In this paper we create, for the first time, a computational model for LAEs 
that emulate the kinematic effects of galaxy rotation and outflows in the \lya
spectral line.

Our main conclusions are:

\begin{itemize}
	\item The effect of rotation \vrot in a LAE spectrum is to broaden the 
	\lya line.

	\item The effect of outflows \vout in a LAE spectrum is to increase the peak 
	asymmetry of the \lya line.

	\item The effect of optical depth \tauh in a LAE spectrum is to induce a 
	redshift around the zero velocity in the \lya line.

	\item Rotation induces a dependency in viewing angle on the \lya line. The 
	closer the observer is to the equator of the galaxy, the higher is the central 
	valley of the frequency distribution.

	\item Rotations induces a Doppler Shift in frequency of the only-outflows 
	spectrum of the LAE.

	\item The final spectra obtained are roughly consistent with LAEs 
	observations. 
\end{itemize}

@@

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation_small}
	\end{center}
	\caption{\textbf{Caption:} $\vout=50$\kms.
		\label{fig:varying_rotation_small}}
\end{figure*}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow_small}
	\end{center}
	\caption{\textbf{Caption:} $\vrot=50$\kms.
		\label{fig:varying_outflow_small}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation}
	\end{center}
	\caption{\textbf{Caption:} $\vout=50$\kms.
		\label{fig:varying_rotation}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow}
	\end{center}
	\caption{\textbf{Caption:} $\vrot=50$\kms.
		\label{fig:varying_outflow}}
\end{figure*}

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We find three important consequences of rotation.
First, it increases the detected flux at the line's center in comparison with the
non-rotational case, second it produces a line broadening and third it introduces
a dependency with the  viewing angle.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the main features of the
Monte Carlo simulations.
As an application we constraint the rotation parameters of a set of observed LAE
emission spectra.
\end{abstract}

@@

In this paper we explore, for the first time in the literature,
the results of a model for the emergent \lya\ line from rotating outflows.
The main results for the model are computed from a Monte-Carlo radiative transfer
simulation and confronted with a simple semi-analytic ansatz that adds the effects
of rotation onto results of pure outflow kinematics.

The main effects of rotation on the \lya\ line morphology are:

\begin{itemize}
  \item Broadening the line.
	\item Increasing the intensity at the line's center.
	\item Inducing a dependency on the viewing angle,
  The closer the observer is to the pole of the galaxy (defined by the rotation axis),
  the smaller are the effects from rotation.
\end{itemize}

We found that in the case of solid body rotation these effects can be quantitatively
explained by a Doppler boost, where the Doppler factor can be computed as the
 producet of quantities at the surface of last scattering, namely $\vec{v}_{\rm rot}\cdot \hat{k}$,
 where $\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is the direction
 of the photon's propagation.
These conclusions trenghten the evidence reported by \cite{Garavito14} when they modeled
the effect of pure rotation on \lya\ spectra.

@@

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation_small}
	\end{center}
	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
		\label{fig:varying_rotation_small}}
\end{figure*}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow_small}
	\end{center}
	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
		\label{fig:varying_outflow_small}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100,
% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100,
% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100,
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation}
	\end{center}
	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
		\label{fig:varying_rotation}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75,
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow}
	\end{center}
	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
		\label{fig:varying_outflow}}
\end{figure*}
